---
title: "Ecol_Pop_Exam_1_parte_1"
author: "BIOL4558"
date: "Agosto 2021"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Las preguntas del examen de esta parte proviene del capiyulo 1:


1. In 1993, when the first edition of this book was written, the world's human population was expected to double in size in approximately 50 years. Assuming population growth is continuous, calculate r for the human population. 

```{r}
#t_double=ln((2)/r)

#50=ln(2)/r

r=log(2)/50
r
```


2. If the population size in 1993 was 5.4 billion, what was the projected population size for the year 2021?


3.  The future is here!. On August  2, 2000 the best estimate of the world population size was 6.087 billion - different than that projected by the model in 1993. To find out the current estimate of the world populaiton size, visit the website maintained by the US Census Bureau.

[Census Population size website](https://www.census.gov/popclock/)

4. What is today's date and how large is the population of the world today. By what percentage has it increased?

5. Considering the present population size what is the observed growth rate of human population on the planet **per year**, use the population size of 1993 (5.538 billion). 

```{r eval=FALSE, include=FALSE}
r= (7.796-5.538)/(2021-1993)
r
```



6. 




