---
title: "Bienvenido a BIOL4458"
author: "Ecología de Población Aplicada"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


**Instructor:** Raymond L. Tremblay  
 
**Website**: [https://github.com/RaymondLTremblay?tab=repositories](https://raymondltremblay.github.io/Ecologia_de_Poblacion_UPRH/) 	
**Email**: raymondtremblay_at_upr.edu
 
### Horario de la Clase
**Lectures**: L, W a las 11:00am hasta las 12:20 pm (120 mins)  in NL204.     


**Hora de Consulta**:   

- Tremblay: TBD    
 
  

### Libros  
- [Gotelli, N. J. A primer of ecology](https://www.amazon.com/Primer-Ecology-Fourth-Nicholas-Gotelli/dp/0878933182)      
- [Beyond Connecting the Dots](http://beyondconnectingthedots.com/) (free download)   
- Se asignarán lecturas adicionales de la literatura primaria para discusión periódicamente.    

### Software   
- [InsightMaker- free web-based systems modeling tool](https://insightmaker.com/)(no installation needed)        
- [R- free statistical programming language](https://cran.r-project.org/)      
- MS Excel 

### Prerequisitos    
- BIOL 3115-3116 Ecología General 


### Descripción de la cLase 
Esta clase explorará cómo se pueden usar los conceptos de ecología de poblaciones para informar la conservación y el manejo de poblaciones naturales y ecosistemas. Enfatizaremos enfoques prácticos para la resolución de problemas en ecología, conservación y manejo de la vida silvestre a través de la aplicación creativa de la teoría de la ecología de poblaciones utilizando modelos de simulación y estadísticas. Los temas incluirán análisis de viabilidad de la población (PVA), modelos de sostenibilidad de hábitat, modelos de metapoblación, modelos de interacción de especies, amenazas a la viabilidad de la población, gestión de la población de vida silvestre y más. Los ejercicios de laboratorio proporcionarán a los estudiantes experiencia práctica con modelos ecológicos y sus aplicaciones prácticas en la conservación y gestión de poblaciones silvestres.

### Objetivos de aprendizaje
Los estudiantes podrán:

1. Identificar las principales clases de modelos utilizados por los ecologistas (por ejemplo, estadísticos o mecanicistas, cuantitativos o heurísticos, estocásticos o deterministas) y explicar cómo y por qué los ecologistas utilizan estos modelos.
2. Aplicar herramientas como el análisis de viabilidad poblacional (PVA) y modelos de metapoblación para abordar la conservación y manejo de poblaciones naturales.
3. Realice estadísticas básicas, visualización de datos, modelado de simulación y validación de modelos con Excel, el lenguaje de cálculo estadístico 'R' y el software basado en web, InsightMaker.
4. Evaluar críticamente la fuerza de las inferencias extraídas de los modelos de simulación ecológica utilizando herramientas como el análisis de sensibilidad.
5. Explique cómo las interacciones entre especies pueden influir en la dinámica de la población (por ejemplo, predicciones de cambios en el rango de especies).
6. Comunicar la investigación original en población aplicada y ecología comunitaria a través de presentaciones orales y escritas de estilo profesional.   

### Puntación: 
The course grade will be based on the following components:

Lab exercises Vea prontuartio correspondiente         33%   

Cuestionarios y participación                         10% (40 puntos)
Proyecto grupal de un tema                            25% (100 puntos)
Examen parcial n.°1 (fecha por determinar)            10% (40 puntos)
Trabajo de PVA (fecha por determinar)                 10% (40 puntos)
Examen final (TBD)                                    25% (100 puntos)

   
### Examenes:
Habrá un examen parcial y un examen final (acumulativo). Estos consistirán en preguntas de opción múltiple, de respuesta corta y preguntas de ensayo que requieren síntesis de ideas y pensamiento crítico. El exámen parcial y final serán acumulativos y se basarán en toda la información presentada durante la semanas anterior al examen.

### Lectures
Las calificaciones de las conferencias se basarán principalmente en la participación y en las pruebas breves. La participación es esencial para el proceso de aprendizaje (y para nuestro mutuo disfrute de esta clase). El aprendizaje no es un proceso pasivo; Se espera que los estudiantes se involucren con el material en clase en lugar de simplemente escuchar y tomar notas. Debe estar preparado en clase para hacer preguntas, responder preguntas de otros estudiantes y participar en actividades de resolución de problemas en clase.   

### Labs  
Vea el Prontuario

### Proyecto de PVA
Los estudiantes trabajarán en grupos de ~ 2 para realizar un análisis de viabilidad de la población (PVA) para clasificar las acciones de conservación o manejo para una especie de interés de conservación (¡especie de su elección!). La calificación se basará en los productos terminados (presentaciones escritas y orales), así como en la participación y evaluaciones de pares.

### Proyecto de grupal de un tema



### Política de recuperación y trabajo tardío:
Los exámenes y laboratorios perdidos no se pueden recuperar, excepto en el caso de emergencias. Si se pierde una reunión de la clase, es su responsabilidad hablar con uno de sus compañeros de clase sobre lo que se perdió. Si se pierde una reunión de laboratorio, aún es responsable de completar las actividades de laboratorio y escribir en su propio tiempo. Siempre que sea posible, avíseme con anticipación si va a faltar a clases o al laboratorio.



### Students with Disabilities
Any student with a disability needing academic adjustments or accommodations is requested to speak with the Disability Resource Center (Thompson Building, Suite 101) as soon as possible to arrange for appropriate accommodations.

### Statement on Academic Dishonesty
Cheating, plagiarism or otherwise obtaining grades under false pretenses constitute academic dishonesty according to the code of this university. Plagiarism is using the ideas or words of another person without giving credit to the original source; this includes copying another student in class. Always cite the source of your information. This includes copying or paraphrasing from a book, journal, or unpublished material without giving credit to the author(s), and submitting a term paper that was used in another course. Academic dishonesty will not be tolerated and penalties can include filing a final grade of "F"; reducing the student's final course grade one or two full grade points; awarding a failing mark on the coursework in question; or requiring the student to retake or resubmit the coursework. For more details, see the [University of Nevada, Reno General Catalog](http://catalog.unr.edu/).

### Statement on Audio and Video Recording
Surreptitious or covert video-taping of class or unauthorized audio recording of class is prohibited by law and by Board of Regents policy. This class may be videotaped or audio recorded only with the written permission of the instructor. In order to accommodate students with disabilities, some students may have been given permission to record class lectures and discussions. Therefore, students should understand that their comments during class may be recorded.

### Statement for Academic Success Services
Your student fees cover usage of the University Math Center [(775) 784-4433], University Tutoring Center [(775) 784-6801], and [University Writing Center (775) 784-6030]. These centers support your classroom learning; it is your responsibility to take advantage of their services.

### This is a safe space
The University of Nevada, Reno is committed to providing a safe learning and work environment for all. If you believe you have experienced discrimination, sexual harassment, sexual assault, domestic/dating violence, or stalking, whether on or off campus, or need information related to immigration concerns, please contact the University's Equal Opportunity & Title IX Office at 775-784-1547. Resources and interim measures are available to assist you. For more information, please visit: http://www.unr.edu/equal-opportunity-title-ix.

### Statement on COVID-19 Training Policies
Students must complete and follow all guidelines as stated in the Student COVID-19 Training modules, or any other trainings or directives provided by the University.

### Statement on COVID-19 Face Coverings
In response to COVID-19, and in alignment with State of Nevada Governor Executive Orders, Roadmap to Recovery for Nevada plans, Nevada System of Higher Education directives, the University of Nevada President directives, and local, state, and national health official guidelines face coverings are required at all times while on campus, except when alone in a private office. This includes the classroom, laboratory, studio, creative space, or any type of in-person instructional activity, and public spaces.

A “face covering” is defined as a “covering that fully covers a person’s nose and mouth, including without limitation, cloth face mask, surgical mask, towels, scarves, and bandanas” (State of Nevada Emergency Directive 024).

Students that cannot wear a face covering due to a medical condition or disability, or who are unable to remove a mask without assistance may seek an accommodation through the Disability Resource Center.

### Statement on COVID-19 Social Distancing:
Face coverings are not a substitute for social distancing. Students shall observe current social distancing guidelines where possible in accordance with the Phase we are in while in the classroom, laboratory, studio, creative space (hereafter referred to as instructional space) setting and in public spaces. Students should avoid congregating around instructional space entrances before or after class sessions. If the instructional space has designated entrance and exit doors students are required to use them. Students should exit the instructional space immediately after the end of instruction to help ensure social distancing and allow for the persons attending the next scheduled class session to enter.

### Statement on COVID-19 Disinfecting Your Learning Space
Disinfecting supplies are provided for you to disinfect your learning space. You may also use your own disinfecting supplies.

### Statement on COVID-19, COVID-19 Like Symptoms, and Contact with Someone Testing Positive for COVID-19
Students must conduct daily health checks in accordance with CDC guidelines. Students testing positive for COVID 19, exhibiting COVID 19 symptoms or who have been in direct contact with someone testing positive for COVID 19 will not be allowed to attend in-person instructional activities and must leave the venue immediately. Students should contact the Student Health Center or their health care provider to receive care and who can provide the latest direction on quarantine and self-isolation. Contact your instructor immediately to make instructional and learning arrangements.

### Statement on Failure to Comply with Policy (including as outlined in this Syllabus) or Directives of a University Employee
In accordance with section 6,502 of the University Administrative Manual, a student may receive academic and disciplinary sanctions for failure to comply with policy, including this syllabus, for failure to comply with the directions of a University Official, for disruptive behavior in the classroom, or any other prohibited action. “Disruptive behavior" is defined in part as behavior, including but not limited to failure to follow course, laboratory or safety rules, or endangering the health of others. A student may be dropped from class at any time for misconduct or disruptive behavior in the classroom upon recommendation of the instructor and with approval of the college dean. A student may also receive disciplinary sanctions through the Office of Student Conduct for misconduct or disruptive behavior, including endangering the health of others, in the classroom. The student shall not receive a refund for course fees or tuition.




```{r echo=FALSE, eval=FALSE}
rmarkdown::render('index.Rmd', 'word_document')
rmarkdown::render('schedule.Rmd', 'word_document')

knitr::opts_chunk$set(echo = TRUE, message = FALSE)


rmarkdown::render('LAB1.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LAB2.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LAB3.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LAB4.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LAB5.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LAB6.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LAB7.Rmd',rmarkdown::pdf_document())

rmarkdown::render('index.Rmd',rmarkdown::pdf_document())
rmarkdown::render('schedule.Rmd',rmarkdown::pdf_document())
rmarkdown::render('INTRO.Rmd',rmarkdown::pdf_document())
rmarkdown::render('FINAL_PROJECTS.Rmd',rmarkdown::pdf_document())
rmarkdown::render('EXTRA_CREDIT.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE1.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE2.Rmd',rmarkdown::pdf_document())  # latex_engine="xelatex"
rmarkdown::render('LECTURE3.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE4.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE5.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE6.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE7.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE8.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE9.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE10.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE11.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE12.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE13.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE14.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE15.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE16.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE17.Rmd',rmarkdown::pdf_document())
rmarkdown::render('LECTURE18.Rmd',rmarkdown::pdf_document())


rmarkdown::render('PVA1_421.Rmd',rmarkdown::pdf_document())
rmarkdown::render('PVA2_421.Rmd',rmarkdown::pdf_document())
rmarkdown::render('PVA3_421.Rmd',rmarkdown::pdf_document())


## [Quiz #1](https://unr.instructure.com/courses/15896/quizzes/21572)
## [Quiz #2](https://unr.instructure.com/courses/15896/quizzes/24149)

### Field Trips
# We will be taking 1-2 field trips to collect wildlife population data. One field trip will be full-day, on a weekend: bring breakfast, lunch, and plenty of water. Bring any appropriate gear for spending the day in the field including: binoculars, spotting scopes, and field guides, and wear appropriate clothing for hiking. Be prepared for inclement weather.


## UPCOMING

# - metapopulations
# - source sink


        # - text: "Lab 8: mule deer PVA"
        #   href: Lab8_muledeer.docx
        # - text: "Lab 8: mule deer MARK file"
        #   href: PDH_KnownFate.zip


```






```{r echo=FALSE}
rmd2rscript <- function(infile="LECTURE2.Rmd"){    # function for converting markdown to scripts
  outfile <- gsub(".Rmd",".R",infile)
  close( file( outfile, open="w" ) )   # clear output file
  con1 <- file(infile,open="r")
  con2 <- file(outfile,"w")
  stringToFind <- "```{r*"
  stringToFind2 <- "echo"
  isrblock <- FALSE
  #count=0
  blocknum=0
  
  while(length(input <- readLines(con1, n=1)) > 0){   # while there are still lines to be read
    isrblock <- grepl(input, pattern = stringToFind, perl = TRUE)   # is it the start of an R block?
    showit <- !grepl(input, pattern = stringToFind2, perl = TRUE)   # is it hidden (echo=FALSE)
    if(isrblock){
      blocknum=blocknum+1
      while(!grepl(newline<-readLines(con1, n=1),pattern="```",perl=TRUE)){
        if((blocknum>1)&((showit)|(blocknum==2))) write(newline,file=con2,append=TRUE)
        #count=count+1
      }
      isrblock=FALSE
    }
  }
  closeAllConnections()
}

rmd2rscript_labanswers <- function(infile="LECTURE2.Rmd"){    # function for converting markdown to scripts
  outfile <- gsub(".Rmd",".R",infile)
  close( file( outfile, open="w" ) )   # clear output file
  con1 <- file(infile,open="r")
  con2 <- file(outfile,"w")
  stringToFind <- "```{r*"
  stringToFind2 <- c("answer","test")
  isrblock <- FALSE
  #count=0
  blocknum=0
  
  while(length(input <- readLines(con1, n=1)) > 0){   # while there are still lines to be read
    isrblock <- grepl(input, pattern = stringToFind, perl = TRUE)   # is it the start of an R block?
    showit <- grepl(input, pattern = stringToFind2[1], perl = TRUE) | grepl(input, pattern = stringToFind2[2])   
    if(isrblock){
      blocknum=blocknum+1
      while(!grepl(newline<-readLines(con1, n=1),pattern="```",perl=TRUE)){
        if((blocknum>1)&((showit)|(blocknum==2))) write(newline,file=con2,append=TRUE)
        #count=count+1
      }
      isrblock=FALSE
    }
  }
  closeAllConnections()
}

```


```{r echo=FALSE, eval=FALSE}


rmd2rscript("LECTURE7.Rmd")  ##
rmd2rscript("LECTURE8.Rmd")
rmd2rscript("LECTURE15.Rmd")
rmd2rscript("LECTURE16.Rmd")
rmd2rscript("LECTURE17.Rmd")
rmd2rscript("LAB4.Rmd")  ##
rmd2rscript("LAB5.Rmd")
rmd2rscript("LAB7.Rmd")

rmd2rscript("LECTURE12.Rmd")

rmd2rscript("LECTURE9.Rmd")
rmd2rscript("LECTURE10.Rmd")
rmd2rscript("FigureDemo.Rmd")

rmd2rscript("PVA1_421.Rmd")
rmd2rscript("PVA3_421.Rmd")

rmd2rscript("TimeSeries.Rmd")
rmd2rscript("SEM.Rmd")
rmd2rscript("SpatialAutocorrelation.Rmd")
rmd2rscript("BayesianNetworks.Rmd")
rmd2rscript("MixedEffects.Rmd")
rmd2rscript("QuantileRegression.Rmd")
rmd2rscript("Ordination.Rmd")
rmd2rscript("RSFs.Rmd")

rmd2rscript_labanswers("LAB1.Rmd")
rmd2rscript_labanswers("LAB2.Rmd")
rmd2rscript_labanswers("LAB3.Rmd")
rmd2rscript_labanswers("LAB4.Rmd")

```



