---
title: "Tillandsia brachycaulos"
author: "BIOL4558"
date: "Agosto 2021"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### APPENDIX S1
 - R code for figures
 - This appendix provides R code for producing the figures included in the main manuscript.
 - Install the library using install.packages() or by downloading from http://cran.R-project.org


### Load the library ( some minor changes)


```{r}
library(popdemo)
library(popbio)
```


## Lectura

Deben leer el siguiente manuscrito para lunes, habrá prueba corta sobre este manuscrito. 

[Mondragón, Durán, Ramírez, Valverde](Guest_Speakers/Temporal_variation_in_the_demography_of.pdf). Habrá un quiz de comprobación de lectura al principio de la clase.  


```{r}
 Tis=matrix(c(
0,0,0,0.007,0.017,0.036,0,.001,0.011,0.024,0.035,
0.28,0.167,0,0,0,0,0,0,0,0,0,
0,.33,.189,0,0,0,0,0,0,0,0,
0,.103,.378,.134,0,0,0,0,0,0,0,
0,0,.081,.179,.066,0,0,0,0,0,0,
0,0,0.014,0,0.082,0,0,0,0,0,0,
0,.039,0.108,.448,.754,.940,0,.300,0.570,.985,1.243,
0,0,0,0,0,0,.033,.57,0,0,0,
0,0,0,0,0,0,.148,.183,.035,0,0,
0,0,0,0,0,0,.18,.207,.193,.037,0,
0,0,0,0,0,0,.115,.126,.175,.066,0
), byrow=TRUE, ncol=11)

Tis


```

Draw the life history of the matrix


Use the package **Rage**

```{r}
library(Rage)
```


```{r}
plot_life_cycle(Tis)
```






```{r}
isErgodic(Tis, digits=10, return.eigvec=FALSE)
isIrreducible(Tis)
```


```{r}
eigs <- eigen(Tis)
eigs
lambdamax <- Re(eigs$values[1])
lambdamax
lambda(Tis)
truelambda(Tis)
```

***


## Stable stage Distribution

The relative proportion of individuals represented in each age/stage of an exponentially increasing (or decreasing) population.  Once the populaiton reaches a stable age/stage distribution, these proportion remainconstant through time.  A stage age/stage distribution implies that a population is growing with fixed survivorship *lx* and fecundity *bx* schedules.

```{r}

w <- stable.stage(Tis)
w
```

## Reproductive value, *vx*

Reproductive value is the expected number of offspring that remain to be born to individuals of age/stage **x**. By definition the value of newborns *v0* is 1.0. 

```{r}

v <- reproductive.value(Tis)
v
```


## Elasticities and Sensitivity

Both these tools lets you understand the importance of the effect of growth/persistence of the population, which of the parameters are most relevant.



###  Using the elasticity vs. the sensitivity.  
- Elasticity assesses effect of a proportional change in the values such as lambda: In other words if you were able modify the parameter (increase of decrease survivorship or reproduction) what would be the proportional effect on the growth of the population.  In this case the values of reproduction and survivorhsip and growth are comparible (same scale)

- sensitivities assess the effect of an additive change. Reproduction and survivorship and growth are on different scales. 
            
            
            
```{r}
sens(Tis)
```


```{r}
elas(Tis)
```










## TRANSIENT DYNAMICS

__________________________________________________________________________
#
Fig. 2a: Population-specific transient dynamics
#
This figure is designed to be 3.5" by 3.5"
(Check size of plot window using 'par(din)').

Set margins:
```{r}
par(mar=c(10,4,1,1))
```
#Create 2 population vectors.  One is adult-
#biased, and it amplifies.  One is juvenile-
# biased and it attenuates.

```{r}
n0.amp <- c(1000,1,1,1,1,1,1,1,1,1,1)

n0.att <- c(1,1,1,1,1,1,1,1,1,1,1000)
```

Project these vectors using the project function.
We are standardising the matrix using 'standard.A=T'
to remove effects of asymptotic dynamics. This 
means that the projection has a long-term growth
rate of unity, i.e. it does not grow or decline in 
the long-term.  We also standardise the population
vector to sum to 1 using 'standard.vec=T'.
These standardisations make it easier (and fairer)
to compare between models with different population 
sizes, and different long-term growth rates.

```{r}
pr2.1 <- project(Tis, vector=n0.amp, time=156, 
                 standard.A=T, standard.vec=T)
pr2.2 <- project(Tis, vector=n0.att, time=156, 
                 standard.A=T, standard.vec=T)
```
Plot the amplification projection and label it

```{r}
plot(pr2.1, ylim=c(0.4,1), log="y", cex.axis=0.8)
text(52, pr2.1[51], "amplification", 
     adj=c(1,-0.5), cex=0.8)
#maxamp(Tis, vector=n0.amp)
reac(Tis,vector=n0.amp)
inertia(Tis,vector=n0.amp)
```

Calculate the transient dynamics of the
amplification projection

```{r}
reac <- reac(Tis, vector=n0.amp)
#maxamp <- maxamp(Tis, vector=n0.amp, return.t=T)
upinertia <- inertia(Tis, vector=n0.amp)
```
Add points on the projection for amplification
and label them

```{r}
#points(c(1,maxamp$t,31), c(reac,upinertia), 
#       pch=3, col="red")
#text(1, reac, expression(bar(P)[1]), 
#     adj=c(-0.3,0.5), col="red", cex=0.8)
#text(maxamp$t, maxamp$maxamp, expression(bar(P)[max]), 
 #    adj=c(0.1,-0.5), col="red", cex=0.8)
#text(31, upinertia, expression(bar(P)[infinity]), 
#     adj=c(0.1,-0.5), col="red", cex=0.8)
```
Add in the second projection using the
'lines' command and label it
lines(0:50, pr2.2)
text(52, pr2.2[51], "attenuation", adj=c(1,1), cex=0.8)

Calculate the transient dynamics of the 
attenuation projection
firstep <- firststepatt(Lrub1, vector=n0.att)
maxatt <- maxatt(Lrub1, vector=n0.att, return.t=T)
lowinertia <- inertia(Lrub1, vector=n0.att)

firststepatt(Lrub1, vector=n0.att)
maxatt(Lrub1, vector=n0.att, return.t=T)
inertia(Lrub1, vector=n0.att)

Add points on the attenuation projection
and label them
points(c(1,maxatt$t,31), c(firstep,maxatt$maxatt,lowinertia), 
       pch=3, col="red")
text(1, firstep, expression(underline(P)[1]), 
     adj=c(0,-0.6), col="red", cex=0.8)
text(maxatt$t, maxatt$maxatt, expression(underline(P)[min]), 
     adj=c(0.1,1.5), col="red", cex=0.8)
text(31, lowinertia, expression(underline(P)[infinity]), 
     adj=c(0.1,1.5), col="red", cex=0.8)

Add in a dotted line at y=1
lines(c(0,50),c(1,1),lty=2)


#########################################################################

FIGURE 4: TRANSFER FUNCTION MATRIX

This figure is designed to be 7" by 7"
(Check size of plot window using 'par(din)').

This is a nice, easy plot because the function does it all for you.
The functions 'tfmat' and 'inertia.tfmat' calculate a transfer function
for every PPM element.  They are saved as arrays of values.  The functions
are customisable: see '?tfmat' and '?inertia.tfmat'.

Create the array for inertia, using a specific population vector:
n0 <- c(48, 16, 12, 10)
tfmat <- inertia.tfamatrix(Lrub1, vector=n0)

...and plot it!
plot(tfmat)
```{r}
n0 <- c(48, 16, 12, 10)
tfmat <- tfamatrix(Tis)

#...and plot it!
plot(tfmat)
```