---
title: "Tillandsia brachycaulos"
author: "RLT"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### APPENDIX S1
 - R code for figures
 - This appendix provides R code for producing the figures included in the main manuscript.
 - Install the library using install.packages() or by downloading from http://cran.R-project.org


### Load the following libraries


```{r}
library(popdemo)
library(popbio)
```


## Lectura

Deben leer el siguiente manuscrito para lunes, habrá prueba corta sobre este manuscrito. 

[Mondragón, Durán, Ramírez, Valverde](Guest_Speakers/Temporal_variation_in_the_demography_of.pdf). Habrá un quiz de comprobación de lectura al principio de la clase.  


```{r}
 Tis=matrix(c(
0,0,0,0.007,0.017,0.036,0,.001,0.011,0.024,0.035,
0.28,0.167,0,0,0,0,0,0,0,0,0,
0,.33,.189,0,0,0,0,0,0,0,0,
0,.103,.378,.134,0,0,0,0,0,0,0,
0,0,.081,.179,.066,0,0,0,0,0,0,
0,0,0.014,0,0.082,0,0,0,0,0,0,
0,.039,0.108,.448,.754,.940,0,.300,0.570,.985,1.243,
0,0,0,0,0,0,.033,.57,0,0,0,
0,0,0,0,0,0,.148,.183,.035,0,0,
0,0,0,0,0,0,.18,.207,.193,.037,0,
0,0,0,0,0,0,.115,.126,.175,.066,0
), byrow=TRUE, ncol=11)

Tis


```


```{r}
isErgodic(Tis, digits=10, return.eigvec=FALSE)
isIrreducible(Tis)
```


```{r}
eigs <- eigen(Tis)
eigs
lambdamax <- Re(eigs$values[1])
lambdamax
lambda(Tis)
truelambda(Tis)
```


```{r}

w <- stable.stage(Tis)
w
```


```{r}

v <- reproductive.value(Tis)
v
```


## TRANSIENT DYNAMICS

__________________________________________________________________________
#
Fig. 2a: Population-specific transient dynamics
#
This figure is designed to be 3.5" by 3.5"
(Check size of plot window using 'par(din)').

Set margins:
```{r}
par(mar=c(10,4,1,1))
```
#Create 2 population vectors.  One is adult-
#biased, and it amplifies.  One is juvenile-
# biased and it attenuates.

```{r}
n0.amp <- c(1000,1,1,1,1,1,1,1,1,1,1)

n0.att <- c(1,1,1,1,1,1,1,1,1,1,1000)
```

Project these vectors using the project function.
We are standardising the matrix using 'standard.A=T'
to remove effects of asymptotic dynamics. This 
means that the projection has a long-term growth
rate of unity, i.e. it does not grow or decline in 
the long-term.  We also standardise the population
vector to sum to 1 using 'standard.vec=T'.
These standardisations make it easier (and fairer)
to compare between models with different population 
sizes, and different long-term growth rates.

```{r}
pr2.1 <- project(Tis, vector=n0.amp, time=156, 
                 standard.A=T, standard.vec=T)
pr2.2 <- project(Tis, vector=n0.att, time=156, 
                 standard.A=T, standard.vec=T)
```
Plot the amplification projection and label it

```{r}
plot(pr2.1, ylim=c(0.4,1), log="y", cex.axis=0.8)
text(52, pr2.1[51], "amplification", 
     adj=c(1,-0.5), cex=0.8)
#maxamp(Tis, vector=n0.amp)
reac(Tis,vector=n0.amp)
inertia(Tis,vector=n0.amp)
```

Calculate the transient dynamics of the
amplification projection

```{r}
reac <- reac(Tis, vector=n0.amp)
#maxamp <- maxamp(Tis, vector=n0.amp, return.t=T)
upinertia <- inertia(Tis, vector=n0.amp)
```
Add points on the projection for amplification
and label them

```{r}
#points(c(1,maxamp$t,31), c(reac,upinertia), 
#       pch=3, col="red")
#text(1, reac, expression(bar(P)[1]), 
#     adj=c(-0.3,0.5), col="red", cex=0.8)
#text(maxamp$t, maxamp$maxamp, expression(bar(P)[max]), 
 #    adj=c(0.1,-0.5), col="red", cex=0.8)
#text(31, upinertia, expression(bar(P)[infinity]), 
#     adj=c(0.1,-0.5), col="red", cex=0.8)
```
Add in the second projection using the
'lines' command and label it
lines(0:50, pr2.2)
text(52, pr2.2[51], "attenuation", adj=c(1,1), cex=0.8)

Calculate the transient dynamics of the 
attenuation projection
firstep <- firststepatt(Lrub1, vector=n0.att)
maxatt <- maxatt(Lrub1, vector=n0.att, return.t=T)
lowinertia <- inertia(Lrub1, vector=n0.att)

firststepatt(Lrub1, vector=n0.att)
maxatt(Lrub1, vector=n0.att, return.t=T)
inertia(Lrub1, vector=n0.att)

Add points on the attenuation projection
and label them
points(c(1,maxatt$t,31), c(firstep,maxatt$maxatt,lowinertia), 
       pch=3, col="red")
text(1, firstep, expression(underline(P)[1]), 
     adj=c(0,-0.6), col="red", cex=0.8)
text(maxatt$t, maxatt$maxatt, expression(underline(P)[min]), 
     adj=c(0.1,1.5), col="red", cex=0.8)
text(31, lowinertia, expression(underline(P)[infinity]), 
     adj=c(0.1,1.5), col="red", cex=0.8)

Add in a dotted line at y=1
lines(c(0,50),c(1,1),lty=2)


#########################################################################

FIGURE 4: TRANSFER FUNCTION MATRIX

This figure is designed to be 7" by 7"
(Check size of plot window using 'par(din)').

This is a nice, easy plot because the function does it all for you.
The functions 'tfmat' and 'inertia.tfmat' calculate a transfer function
for every PPM element.  They are saved as arrays of values.  The functions
are customisable: see '?tfmat' and '?inertia.tfmat'.

Create the array for inertia, using a specific population vector:
n0 <- c(48, 16, 12, 10)
tfmat <- inertia.tfamatrix(Lrub1, vector=n0)

...and plot it!
plot(tfmat)
```{r}
n0 <- c(48, 16, 12, 10)
tfmat <- tfamatrix(Tis)

#...and plot it!
plot(tfmat)
```