---
title: "Vignette_Tort"
author: "RLTremblay"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, eval=TRUE}
# Install dependencies from CRAN:
#install.packages(c("devtools", "expm", "MCMCpack", "markovchain"))

# Install stable version from GitHub (recommended):
# NOTE don't forget to change the version number!
# devtools::install_github("iainmstott/popdemo/x.x-x/popdemo") #x.x-x is the desired version number
library(popdemo)

# Install development version 'popdemoDev' (not recommended):
#devtools::install_github("iainmstott/popdemo/Dev/popdemoDev", ref = "development")

#library(popdemoDev)

library(popbio)
```



```{r}
data(Tort)

Tort
```

That is, the population size at time t nt is equal to projection matrix A multiplied by the population size at time 0, n0. We have the matrix, but not a vector. We will:
- choose a vector using a random uniform distribution
- project this vector using the project function from popdemo
- plot the projection

```{r}

set.seed(1234)
Tortvec1 <- runif(8) # el 8 se refiere que hay 8 etapas
Tortvec1 <- Tortvec1/sum(Tortvec1) #scales the vector to sum to 1
( Tortp1.1 <- project(Tort, Tortvec1, time = 11) )

plot(Tortp1.1, log = "y")
```





```{r}
#nT=c(10000, 500, 300, .....)
stable.stage(Tort)
vec(Tortp1.1)[1:11, ]
```


The time series of stage 2 sizes is:


```{r}
Tort
vec(Tortp1.1)[, 2]
```


Cuando se usa la funci贸n **eigs** con la opci贸n **"all"** tenemos tres informaci贸n

 - el crecimeinto intrinsico, lambda
 - la etapa estable de distibuci贸n; ss
 - y el valor reproductivo; rv


```{r}
Tort
eigs(Tort, "all")
```
# Start here


```{r}
Tortw <- eigs(Tort, "ss")
Tortpw <- project(Tort, Tortw, time = 100)
```



```{r}
Tortp1.1s <- project(Tort, Tortvec1, time = 100, 
                     standard.A = TRUE, standard.vec = TRUE)
Tortpws <- project(Tort, Tortw, time = 100, 
                   standard.A = TRUE, standard.vec = TRUE)
plot(Tortp1.1s, log = "y")
lines(Tortpws, lty = 2)
```



```{r}
( r1 <- reac(Tort, Tortvec1) )
##   [1] 1.136

( i1 <- inertia(Tort, Tortvec1) )
##   [1] 1.777

## Copiar el chunk ariba
Tortp1.1s <- project(Tort, Tortvec1, time = 100, 
                     standard.A = TRUE, standard.vec = TRUE)
Tortpws <- project(Tort, Tortw, time = 100, 
                   standard.A = TRUE, standard.vec = TRUE)
plot(Tortp1.1s, log = "y")
lines(Tortpws, lty = 2)
points(c(1, 100), c(r1, i1), pch = 3, col = "red")
```




```{r}
TortAMP <- c(1, 1, 2, 3, 5, 8, 13, 21) #a population that amplifies
TortATT <- c(21, 13, 8, 5, 3, 2, 1, 1) #a population that attenuates
TortBTH <- c(0, 0, 0, 1, 0, 0, 0, 0) #a population that does both
Tortvec3 <- cbind(AMP = TortAMP, 
                  ATT = TortATT,
                  BTH = TortBTH)
Tortp3.1 <- project(Tort, Tortvec3, time = 100, 
                    standard.A = TRUE, standard.vec = TRUE)
plot(Tortp3.1, log = "y"); lines(Tortpws, lty = 2)
( r3 <- apply(Tortvec3, 2, reac, A = Tort) )
##      AMP    ATT    BTH 
##   2.6319 0.9153 0.8757
( r3t <- rep(1, 3) )
##   [1] 1 1 1

( i3 <- apply(Tortvec3, 2, inertia, A = Tort) )
##      AMP    ATT    BTH 
##   4.1442 0.9123 1.8019
( i3t <- rep(100, 3) )
##   [1] 100 100 100

( max3 <- apply(Tortvec3[,c(1,3)], 2, maxamp, A = Tort) )
##     AMP   BTH 
##   5.111 1.962
( max3t <- apply(Tortvec3[,c(1,3)], 2, function(x){
                 maxamp(vector = x, A = Tort, return.t = TRUE)$t}) )
##   AMP BTH 
##     6  13

( min3 <- apply(Tortvec3[,c(2,3)], 2, maxatt, A = Tort) )
##      ATT    BTH 
##   0.8377 0.7261
( min3t <- apply(Tortvec3[,c(2,3)], 2, function(x){
                 maxatt(vector = x, A = Tort, return.t = TRUE )$t}) )
##   ATT BTH 
##     4   3

points(c(r3t, i3t, max3t, min3t), 
       c(r3, i3, max3, min3), 
       pch = 3, col = "red")
```



```{r}
plot(Tortp1.1s, log = "y", bounds = TRUE)
```


```{r}
plot(project(Tort, standard.A = TRUE), log = "y")
```


```{r}
Tortpd <- project(Tort, "diri", time = 31,
                  standard.A = TRUE)
##   Warning in .recacheSubclasses(def@className, def, env): undefined subclass
##   "numericVector" of class "Mnumeric"; definition not updated
plot(Tortpd, plottype = "shady", bounds = T, log = "y")
```





